Enhancing the sales prediction of newly published books by means of computational intelligence methods

### 2016-02-02 paper.tex
Problem description
Revised, fixed some things and new text added ;)

@JJ please enhance Table 1

%********************************************************************************
\section{The problem}
\label{sec:problem}

The main issue of this paper is to forecast sales for 
unpublished books for which no historical data is available, just some
descriptive variables relative to the new book. In order to do that,
training data includes historical sales data for all books published
in Spain for several years, including those descriptive variables.

%%% Principio de párrafos a eliminar o mover a estado del arte.
%% Fin de párrafos a eliminar.
  % [pedro] eliminados

The initial dataset, provided in the framework of a contract by
Trevenque Editorial, included 3159 books, and included descriptive information 
for each book, such as price, points of sale, subject, publisher, sales, returns, 
among other information.
Table \ref{tabla:paramsOrig50} shows the original set of parameters describing 
each book.

\begin{table}[!ht]
\caption{Parameters describing a book, as provided by the company Trevenque.}
\label{tabla:paramsOrig50}
\begin{center}
\begin{tabular}{|c|c|}
\hline

@@

\begin{minipage}{2.3in} \begin{verbatim}
rest of distributed 1st year
reprints
number of reprints
total returns
returns - dept stores
delegates returns
rest of returns
total returns firt year
returns -  dept stores 1st year
delegates returns 1st year
rest of returns 1st year
gifts
units distributed as novelty
total points of sale
points of sale - dept stores
delegates points of sale
rest of points of sale
total points of sale 1st year
dept stores points of sale 1st year
delegates points of sale 1st year
rest of points of sale 1st year
weeks for sale
positive sales environment
medium sales environment
negative sales environment
\end{verbatim} \end{minipage}    \\

@@

In this prediction problem there are other external variables that also affect 
the sales of a new book, such as those related to seasonality, trends or 
promotions, that classical prediction methods can not deal with.
However, an expert can consider them to modify the forecast obtained by 
the automatic method. Even publishing companies tend to define rules or 
indexes to subjectively adjust the results of prediction based on
external data. %so what?

@@

% Mean absolute error MAE / Root mean squared error RMSE /  Relative absolute error RAE / Root relative squared error RRSE
  % ¿Qué sentido tiene usar los 4 métodos de error? ¿Es que vamos a seleccionar uno? - JJ
  %[pedro] es lo habitual en los trabajos en los que se realizan tareas de predicción
  % en algunos calculan muchos más, pero estos son los más usados. 
\begin{table*}[h]
\caption{Results (errors and time) obtained using the different forecasting
methods on the original dataset (16 variables per book). In this case,
due to the categorial variables, the MLP could not be run.}
\label{tabla:resultsT}
\centering{}%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
Base & Correlation coefficient (\%) & MAE & RMSE & RAE(\%) & RRSE (\%) & Time(sec)\tabularnewline
\hline 
\hline 
M5P & 0.9076 & \textbf{202.1547} & 591.0257 & \textbf{35.31}\% & 42.10\% & 39.73\tabularnewline
\hline 
LR & 0.9204 & 255.0526 & 548.0975 & 44.5527 & 39.10\% & 0.03\tabularnewline
\hline 
SVM & \textbf{0.9402} & 203.4416 & \textbf{481.7528} & 35.54\% & \textbf{34.37}\% & 308.31\tabularnewline
\hline 
kNN & 0.7273 & 310.4078 & 993.2663 & 54.22\% & 70.85\% & 0\tabularnewline
\hline 
MLP & 0.9197 & 289.7393 & 585.6445 & 50.61\% & 41.78\% & 82.08\tabularnewline
\hline 
\end{tabular}
\end{table*}

Obtained results (time to build the model and error) using the different 
forecasting methods on the dataset with 7 variables per book (obtained 
using the SOM analysis) can be shown in Table \ref{tabla:resultsSOM}.
\begin{table*}[h]
\caption{Results (errors and time) obtained using the different forecasting methods
on the dataset obtained using the SOM feature selection method (6 variables
per book).}
\label{tabla:resultsSOM}
\centering{}%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
Base  & Correlation coefficient (\%)  & MAE  & RMSE  & RAE(\%)  & RRSE (\%)  & Time(sec)\tabularnewline
\hline 
\hline 
M5P  & 0.9161 & \textbf{198.8864} & 562.2113 & \textbf{34.74}\% & 40.11\% & 0.76\tabularnewline
\hline 
LR  & 0.9179 & 263.1702 & 556.3372 & 45.97\% & 39.69\% & 0.07\tabularnewline
\hline 
SVM  & \textbf{0.9388} & 207.7159 & \textbf{486.8913} & 36.28\% & \textbf{34.73}\% & 157.87\tabularnewline
\hline 
kNN  & 0.9304 & 202.0939 & 525.4642 & 35.30\% & 37.48\% & 0\tabularnewline
\hline 
MLP  & 0.9193 & 305.78 & 590.8412 & 53.41\% & 42.15\% & 25.83\tabularnewline
\hline 
\end{tabular}
\end{table*}

@@

\begin{table*}[h]
\caption{Results (errors and time) obtained using the different forecasting methods
on the dataset obtained using the ReliefFAttributeEval feature selection method (8
variables per book).}
\label{tabla:results9}
\centering{}%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
RAE & Correlation coefficient (\%) & MAE & RMSE & RAE(\%) & RRSE (\%) & Time(sec)\tabularnewline
\hline 
\hline 
M5P & 0.9131 & \textbf{197.2886} & 580.328 & \textbf{34.46}\% & 41.40\% & 0.74\tabularnewline
\hline 
LR & 0.9209 & 254.3979 & 546.3899 & 44.44\% & 38.98\% & 0.03\tabularnewline
\hline 
SVM & \textbf{0.9394} & 204.2075 & \textbf{484.2617} & 35.67\% & \textbf{34.54}\% & 232.99\tabularnewline
\hline 
kNN & 0.9103 & 222.0827 & 597.8885 & 38.79\% & 42.65\% & 0\tabularnewline
\hline 
MLP & 0.9199 & 296.3581 & 575.9623 & 51.77\% & 41.09\% & 32.52\tabularnewline
\hline 
\end{tabular}
\end{table*}

@@

\begin{table*}[h]
\caption{Results (errors and time) obtained using the different forecasting methods
on the dataset obtained using the CfsSubsetEval feature selection method (4 variables
per book).}
\label{tabla:results4}
\centering{}%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
CFS & Correlation coefficient (\%) & MAE & RMSE & RAE(\%) & RRSE (\%) & Time(sec)\tabularnewline
\hline 
\hline 
M5P & 0.9255 & \textbf{206.6638} & 531.306 & \textbf{36.10}\% & 37.90\% & 0.56\tabularnewline
\hline 
LR & 0.9194 & 259.3791 & 551.2005 & 45.31\% & 39.32\% & 0.05\tabularnewline
\hline 
SVM & \textbf{0.9342} & 213.0254 &\textbf{ 502.4048} & 37.21\% & \textbf{35.84}\% & 70.12\tabularnewline
\hline 
kNN & 0.9285 & 220.6548 & 527.7374 & 38.54\% & 37.65\% & 0\tabularnewline
\hline 
MLP & 0.9047 & 369.3486 & 646.1241 & 64.52\% & 46.09\% & 10.93\tabularnewline
\hline 
\end{tabular}
\end{table*}

>>>>

%********************************************************************************
\section{Problem description}
\label{sec:problem}

The main objective of this paper is to forecast sales for 
unpublished books for which no historical data is available. Instead, just some
descriptive and pre-sales known variables related to the new book. In order to do that, we have used a training dataset which includes historical sales data for a big amount of books published in Spain for several years, with a set of 50 features or variables per pattern.

The initial dataset, provided in the framework of a contract by Trevenque Editorial S.L., included 3159 books. Table \ref{tabla:paramsOrig50} shows the initial set of variables describing each book.

\begin{table}[!ht]
\caption{Parameters describing a book, provided by the company Trevenque S.L.}
\label{tabla:paramsOrig50}
\begin{center}
\begin{tabular}{|c|c|}
\hline

@@

\begin{minipage}{2.3in} \begin{verbatim}
rest of distributed 1st year
reprints
number of reprints
total returns
returns - dept stores
delegates returns
rest of returns
total returns first year
returns -  dept stores 1st year
delegates returns 1st year
rest of returns 1st year
gifts
units distributed as novelty
total points of sale
points of sale - dept stores
delegates points of sale
rest of points of sale
total points of sale 1st year
dept stores points of sale 1st year
delegates points of sale 1st year
rest of points of sale 1st year
weeks for sale
positive sales environment
medium sales environment
negative sales environment
\end{verbatim} \end{minipage}    \\

@@

% Antonio - briefly describe the variables

In this prediction problem there are other external variables that also affect 
the sales of a new book, such as those related to seasonality, trends or 
promotions, that classical prediction methods can not deal with.
However, an expert can consider them to modify the forecast obtained by 
the automatic method. Even publishing companies tend to define rules or 
indexes to subjectively adjust the results of prediction based on
external data. 
Thus, the proposed method, will analyse the provided data and variables and will yield a forecast of sales according to them. This result could be used as a reference for a human expert, who would consider it in his/her own prediction process.

@@

% Mean absolute error MAE / Root mean squared error RMSE /  Relative absolute error RAE / Root relative squared error RRSE
  % ¿Qué sentido tiene usar los 4 métodos de error? ¿Es que vamos a seleccionar uno? - JJ
  %[pedro] es lo habitual en los trabajos en los que se realizan tareas de predicción
  % en algunos calculan muchos más, pero estos son los más usados. 
\begin{table*}[ht]
\caption{Results (errors and time) obtained using the different forecasting
methods on the original dataset (16 variables per book). In this case,
due to the categorial variables, the MLP could not be run.}
\label{tabla:resultsT}
\centering{}%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
Base & Correlation coefficient (\%) & MAE & RMSE & RAE(\%) & RRSE (\%) & Time(sec)\tabularnewline
\hline 
\hline 
M5P & 0.9076 & \textbf{202.1547} & 591.0257 & \textbf{35.31}\% & 42.10\% & 39.73\tabularnewline
\hline 
LR & 0.9204 & 255.0526 & 548.0975 & 44.5527 & 39.10\% & 0.03\tabularnewline
\hline 
SVM & \textbf{0.9402} & 203.4416 & \textbf{481.7528} & 35.54\% & \textbf{34.37}\% & 308.31\tabularnewline
\hline 
kNN & 0.7273 & 310.4078 & 993.2663 & 54.22\% & 70.85\% & 0\tabularnewline
\hline 
MLP & 0.9197 & 289.7393 & 585.6445 & 50.61\% & 41.78\% & 82.08\tabularnewline
\hline 
\end{tabular}
\end{table*}

Obtained results (time to build the model and error) using the different 
forecasting methods on the dataset with 7 variables per book (obtained 
using the SOM analysis) can be shown in Table \ref{tabla:resultsSOM}.
\begin{table*}[ht]
\caption{Results (errors and time) obtained using the different forecasting methods
on the dataset obtained using the SOM feature selection method (6 variables
per book).}
\label{tabla:resultsSOM}
\centering{}%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
Base  & Correlation coefficient (\%)  & MAE  & RMSE  & RAE(\%)  & RRSE (\%)  & Time(sec)\tabularnewline
\hline 
\hline 
M5P  & 0.9161 & \textbf{198.8864} & 562.2113 & \textbf{34.74}\% & 40.11\% & 0.76\tabularnewline
\hline 
LR  & 0.9179 & 263.1702 & 556.3372 & 45.97\% & 39.69\% & 0.07\tabularnewline
\hline 
SVM  & \textbf{0.9388} & 207.7159 & \textbf{486.8913} & 36.28\% & \textbf{34.73}\% & 157.87\tabularnewline
\hline 
kNN  & 0.9304 & 202.0939 & 525.4642 & 35.30\% & 37.48\% & 0\tabularnewline
\hline 
MLP  & 0.9193 & 305.78 & 590.8412 & 53.41\% & 42.15\% & 25.83\tabularnewline
\hline 
\end{tabular}
\end{table*}

@@

\begin{table*}[ht]
\caption{Results (errors and time) obtained using the different forecasting methods
on the dataset obtained using the ReliefFAttributeEval feature selection method (8
variables per book).}
\label{tabla:results9}
\centering{}%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
RAE & Correlation coefficient (\%) & MAE & RMSE & RAE(\%) & RRSE (\%) & Time(sec)\tabularnewline
\hline 
\hline 
M5P & 0.9131 & \textbf{197.2886} & 580.328 & \textbf{34.46}\% & 41.40\% & 0.74\tabularnewline
\hline 
LR & 0.9209 & 254.3979 & 546.3899 & 44.44\% & 38.98\% & 0.03\tabularnewline
\hline 
SVM & \textbf{0.9394} & 204.2075 & \textbf{484.2617} & 35.67\% & \textbf{34.54}\% & 232.99\tabularnewline
\hline 
kNN & 0.9103 & 222.0827 & 597.8885 & 38.79\% & 42.65\% & 0\tabularnewline
\hline 
MLP & 0.9199 & 296.3581 & 575.9623 & 51.77\% & 41.09\% & 32.52\tabularnewline
\hline 
\end{tabular}
\end{table*}

@@

\begin{table*}[ht]
\caption{Results (errors and time) obtained using the different forecasting methods
on the dataset obtained using the CfsSubsetEval feature selection method (4 variables
per book).}
\label{tabla:results4}
\centering{}%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
CFS & Correlation coefficient (\%) & MAE & RMSE & RAE(\%) & RRSE (\%) & Time(sec)\tabularnewline
\hline 
\hline 
M5P & 0.9255 & \textbf{206.6638} & 531.306 & \textbf{36.10}\% & 37.90\% & 0.56\tabularnewline
\hline 
LR & 0.9194 & 259.3791 & 551.2005 & 45.31\% & 39.32\% & 0.05\tabularnewline
\hline 
SVM & \textbf{0.9342} & 213.0254 &\textbf{ 502.4048} & 37.21\% & \textbf{35.84}\% & 70.12\tabularnewline
\hline 
kNN & 0.9285 & 220.6548 & 527.7374 & 38.54\% & 37.65\% & 0\tabularnewline
\hline 
MLP & 0.9047 & 369.3486 & 646.1241 & 64.52\% & 46.09\% & 10.93\tabularnewline
\hline 
\end{tabular}
\end{table*}
